HO unification from object language to meta language

### 2024-03-18 paper/main.tex
more tex

\subsection{The intuition in a nutshell}
A term $s$ is compiled in a term $t$ where any
``problematic'' subterm $tp$ is replaced by a unification variable
$hp$ and an accessory
(suspended) unification problem $(hp \Ue tp) \in \mathcal{L}$.
Since $t \in \llambda$, \Ue is well behaved (captures \Eo).

@@

\begin{definition}[\maybebeta]
  $t$ is in \maybebeta if it contains a subterm matching the
  pattern \elpiIn{app[uva _ _|_]}
\end{definition}

\newcommand{\normal}{\ensuremath{\mathrm{normal}}}
\begin{definition}[Normal form]
Given a set of terms $X \subseteq \Ho{}$,
$$
\normal(X) =
X \cap (\maybebeta{} ~\cup~ \maybeeta{}) = \emptyset
$$
\end{definition}

\noindent
We write $\sigma X = \{ \sigma t_i | t_i \in X\}$.

\begin{proposition}[Normal form preservation]\label{prop:nf}
$\forall \mathcal{T}, \forall \mathbb{L}, \forall p,$
$$
\normal(\sigma\mathcal{T}) \land
\stepH{\mathcal{T}}{p}{\sigma}{\mathbb{L}}{\sigma'}
\Rightarrow \normal(\sigma'\mathcal{T})
$$
\end{proposition}

TODO: check if this is also true for $\rho$ after decomp.
In particular this guaarantees that is we start from normal terms
we never introduce eta-long or non-beta-normal terms. This is required in
order to show the fidelity, since there terms would not be considered
as equal by \Ue.

Problematic terms are terms where the head constructor may
change if instantiated, that is, they are in \maybebeta or \maybeeta.
On these terms \Ue is weak/incorrect.
\begin{itemize}
\item \elpiIn{app[F,con "a"]} may be \Eo to \elpiIn{con "a"}
\item \elpiIn{lam x\lam y\app[F,y,x]} may be \Eo to \elpiIn{con "f"} 
\end{itemize}

\noindent
\solve{} solves links when they are no more ``problematic'', for example
in the examples above if \elpiIn{F} materializes, or if the
other side of the link forces a commit. EXPLAIN.
In these cases the problems are ``resumed'' by using \Ue,
otherwise they are kept suspended until the
decompilation phase.
\solve also detects occur checks trough the suspended unif problems
in order to grant \cref{prop:fidelity}.

Note tha the first case is outside $\llambda$ but is frequent in practice
so we handle it. It is also what prop 5 is about.

Note that proposition \ref{prop:nf} does not hold in this case since
decompilation can introduce terms in \maybebeta or \maybebeta.
\section{Alternative encodings and related work}

>>>>

\subsection{The intuition in a nutshell}
A term $s$ is compiled in a term $t$ where any
``problematic'' subterm $tp$ is replaced by a unification variable $h$
and an accessory link, that represent a suspended unification problem
$h \Ue tp$. As a result \Ue is well behaved on $t$, that is it captures \Eo.
We now define ``problematic'' formally.

@@

\begin{definition}[\maybebeta]
  $\maybebeta = \{ X t_1 \ldots t_n | t \not\in \llambda \}$.
\end{definition}

\noindent
An example of $t$ in \maybebeta{} is $F a$ for a constant $a$. Note however tha
an oracle could provide an assignment $\rho = \{ F \mapsto \lambda x.x\}$
that makes the resulting term fall outside of \maybebeta.

\begin{definition}[Subterm $\mathcal{P}(t)$]
$$
\begin{array}{l}
t \in \mathcal{P}(t)\\
t = f t_1\ldots t_n \Rightarrow \mathcal{P}(t_i) \subseteq \mathcal{P}(t) \land f \in \mathcal{P}(t)\\
t = \lambda x.t' \Rightarrow \mathcal{P}(t') \subseteq \mathcal{P}(t)\\
\end{array}
$$
\end{definition}

\noindent
We write $\mathcal{P}(X) = \bigcup_{t\in X} \mathcal{P}(t)$.

\marginpar{TODO: find better word}
\newcommand{\normal}{\ensuremath{\mathrm{normal}}}
\begin{definition}[Normal form]
Given a set of terms $X \subseteq \Ho{}$,
$$
\normal(X) =
\mathcal{P}(X) \cap (\maybebeta{} ~\cup~ \maybeeta{}) = \emptyset
$$
\end{definition}

\noindent
We write $\sigma X = \{ \sigma t | t \in X\}$.

\begin{proposition}[Normal form preservation]\label{prop:nf}
$\forall \mathcal{T}, \forall \mathbb{L}, \forall p, \forall \sigma, \forall \sigma'$
$$
\normal(\sigma\mathcal{T}) \land
\stepH{\mathcal{T}}{p}{\sigma}{\mathbb{L}}{\sigma'}
\Rightarrow \normal((\sigma \cup \sigma')\mathcal{T})
$$
\end{proposition}

\noindent
In particular this guarantees that is we start from normal terms
we never introduce eta-long or non-beta-normal terms in $\sigma'$.

Note that proposition \ref{prop:nf} does not hold for \Uo{} since
decompilation can introduce (actually restore) terms in
\maybeeta or \maybebeta.

\section{Alternative encodings and related work}
